Structured Versus Decorated Cospans

### 2021-04-06 cospans_new.tex
Second attempt at proving Theorem 2.1

Lastly, the monoidal bifibration $U_1$ satisfies the Beck--Chevalley condition. Given a pushout square in $\A$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1.5,0) {$b$};
\node (C) at (0,-1) {$c$};
\node (D) at (1.5,-1) {$d$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$h$} (B)
(A) edge node[left]{$k$} (C)
(B) edge node[right]{$g$} (D)
(C) edge node[below]{$f$} (D);
\end{tikzpicture}
\]
we need to show that the following square commutes up to a natural isomorphism:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$(\A/U)_a$};
\node (B) at (1.5,0) {$(\A/U)_b$};
\node (C) at (0,-1) {$(\A/U)_c$};
\node (D) at (1.5,-1) {$(\A/U)_d$};
\node (E) at (0.75,-0.5) {$\Searrow$};
\path[->,font=\scriptsize,>=angle 90]
(B) edge node[above]{$h^*$} (A)
(A) edge node[left]{$k_!$} (C)
(B) edge node[right]{$g_!$} (D)
(D) edge node[below]{$f^*$} (C);
\end{tikzpicture}
\]
Let $(b,x,b \xrightarrow{i} U(x))$ be an object of $(\A/U)_b$. Going right, we precompose with $h \colon a \to b$ to obtain the object $(a,x,a \xrightarrow{i \circ h} U(x))$ in $(\A/U)_a$. Then going down, we push forward along $k$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1.5,0) {$U(x)$};
\node (C) at (0,-1) {$c$};
\node (D) at (1.5,-1) {$c+_{a}U(x)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i \circ h$} (B)
(A) edge node[left]{$k$} (C)
(B) edge node[right]{$$} (D)
(C) edge node[below]{$\psi$} (D);
\end{tikzpicture}
\]
to obtain the object $(c,z,c \to U(z))$ of $(\A/U)_c$, where $z \in \int F$ has $U(z)=c+_a U(x)$. In particular, $z=L(c)+_{L(a)}x$ and as $U$ preserves finite colimits and $UL=1_\A$, we have that 
\[  U(z)=U(L(c)+_{L(a)}x)=U(L(c))+_{U(L(a))}U(x)=c+_a U(x).\]
{\john Need to refer to a proposition where we show $U$ preserves pushouts strictly.}
 For the other route, first going down, we push the object $(b,x,b \xrightarrow{i} U(x))$ forward along $g \colon b \to d$ by taking the following pushout:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$b$};
\node (B) at (1.5,0) {$U(x)$};
\node (C) at (0,-1) {$d$};
\node (D) at (1.5,-1) {$d+_{b}U(x)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(A) edge node[left]{$g$} (C)
(B) edge node[right]{$$} (D)
(C) edge node[below]{$\theta$} (D);
\end{tikzpicture}
\]
which yields the object $(d,z',d \to U(z'))$ of $(\A/U)_d$ where $z' \in \int F$ such that $U(z')=d+_b U(x)$. In particular, $z'$ is given by $z'=L(d)+_{L(b)}x$. We then precompose with $f \colon c \to d$ to obtain the object $(c,z',c \to U(z'))$ of $(\A/U)_c$.

Now we seek a natural isomorphism between the two objects $(c,z,c \to c+_a U(x))$ and $(c,z',c \to d+_b U(x))$ in $(\A/U)_c$. It suffices to obtain a natural isomorphism between the two objects $z=L(c)+_{L(a)} x$ and $z'=L(d)+_{L(b)} x$ in $\X$. This pushout square:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(a)$};
\node (B) at (1.5,0) {$x$};
\node (C) at (0,-1) {$L(c)$};
\node (D) at (1.5,-1) {$L(c)+_{L(a)}x$};
%\node (A') at (4,0) {$L(b)$};
%\node (B') at (6,0) {$x$};
%\node (C') at (4,-1) {$L(d)$};
%\node (D') at (6,-1) {$L(d)+_{L(b)}x$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$I \circ L(h)$} (B)
(A) edge node[left]{$L(k)$} (C)
(B) edge node[right]{$$} (D)
(C) edge node[below]{$\psi$} (D);
%(A') edge node[above]{$I$} (B')
%(A') edge node[left]{$L(g)$} (C')
%(B') edge node[right]{$$} (D')
%(C') edge node[below]{$\theta$} (D');
\end{tikzpicture}
\]
exhibits $z$ as a pushout of the span 
\[  L(c) \xleftarrow{L(k)} L(a) \xrightarrow{I \circ L(h)} x 
\]
in $\X$. As $L$ preserves pushouts, we may apply it to the original pushout square in $\A$
and paste on another pushout square:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(b)$};
\node (B) at (1.5,0) {$x$};
\node (C) at (0,-1) {$L(d)$};
\node (D) at (1.5,-1) {$L(d)+_{L(b)}x$};
\node (E) at (-1.5,0) {$L(a)$};
\node (F) at (-1.5,-1) {$L(c)$};
\path[->,font=\scriptsize,>=angle 90]
(E) edge node [left] {$L(k)$} (F)
(E) edge node [above] {$L(h)$} (A)
(F) edge node [below] {$L(f)$} (C)
(A) edge node[above]{$I$} (B)
(A) edge node[left]{$L(g)$} (C)
(B) edge node[right]{$$} (D)
(C) edge node[below]{$\theta$} (D);
\end{tikzpicture}
\]
obtaining a pushout square 
which exhibits $z'$ as a pushout of the same span in $\X$.  The resulting natural isomorphism
between $z$ and $z'$ gives the desired natural isomorphism between our two objects of $(\A/U)_c$, showing that $U_1$ satisfies the Beck--Chevalley condition.
\end{proof}

>>>>

Lastly, the monoidal bifibration $U_1$ satisfies the Beck--Chevalley condition. Given a pushout square in $\A$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1.5,0) {$b$};
\node (C) at (0,-1) {$c$};
\node (D) at (1.5,-1) {$d$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$h$} (B)
(A) edge node[left]{$k$} (C)
(B) edge node[right]{$g$} (D)
(C) edge node[below]{$f$} (D);
\end{tikzpicture}
\]
we need to show that the following square commutes up to a natural isomorphism:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$(\A/U)_a$};
\node (B) at (1.5,0) {$(\A/U)_b$};
\node (C) at (0,-1) {$(\A/U)_c$};
\node (D) at (1.5,-1) {$(\A/U)_d$};
\node (E) at (0.75,-0.5) {$\Searrow$};
\path[->,font=\scriptsize,>=angle 90]
(B) edge node[above]{$h^*$} (A)
(A) edge node[left]{$k_!$} (C)
(B) edge node[right]{$g_!$} (D)
(D) edge node[below]{$f^*$} (C);
\end{tikzpicture}
\]
Let $(b,x,b \xrightarrow{i} U(x))$ be an object of $(\A/U)_b$. Going left, we precompose with $h \colon a \to b$ to obtain the object $(a,x,a \xrightarrow{i \circ h} U(x))$ of $(\A/U)_a$. Then going down, we push forward along $k$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1.5,0) {$U(x)$};
\node (C) at (0,-1) {$c$};
\node (D) at (1.5,-1) {$c+_{a}U(x)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i \circ h$} (B)
(A) edge node[left]{$k$} (C)
(B) edge node[right]{$$} (D)
(C) edge node[below]{$\psi$} (D);
\end{tikzpicture}
\]
to obtain the object $(c,z,c \xrightarrow{\psi} U(z))$ of $(\A/U)_c$, where $z \in \int F$ is given by $z=\bot_c+_{\bot_a}x$. Explicitly, $z$ is given by
\[ 1 \xrightarrow{x} F(U(x)) \xrightarrow{F(\zeta)} F(0+_0 U(x)) \xrightarrow{F(!_c +_{!_a} 1)} F(c+_a U(x)) \]
and so by construction, we have that 
\[  U(z)=U(\bot_c+_{\bot_a}x)=c+_a U(x).\]
{\john Need to refer to a proposition where we show $U$ preserves pushouts strictly.}
 For the other route, first going down, we push the object $(b,x,b \xrightarrow{i} U(x))$ forward along $g \colon b \to d$ by taking the following pushout:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$b$};
\node (B) at (1.5,0) {$U(x)$};
\node (C) at (0,-1) {$d$};
\node (D) at (1.5,-1) {$d+_{b}U(x)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(A) edge node[left]{$g$} (C)
(B) edge node[right]{$$} (D)
(C) edge node[below]{$\theta$} (D);
\end{tikzpicture}
\]
which yields the object $(d,z',d \xrightarrow{\theta} U(z'))$ of $(\A/U)_d$ where $z' \in \int F$ is given by $z'=\bot_d+_{\bot_b}x$. We then precompose with $f \colon c \to d$ to obtain the object $(c,z',c \xrightarrow{\theta \circ f} U(z'))$ of $(\A/U)_c$.

Now we seek a natural isomorphism between the two objects $(c,z,c \xrightarrow{\psi} c+_a U(x))$ and $(c,z',c \xrightarrow{\theta \circ f} d+_b U(x))$ in $(\A/U)_c$. It suffices to obtain a natural isomorphism between the two objects $z=\bot_c+_{\bot_a} x$ and $z'=\bot_d+_{\bot_b} x$ in $\X$. This pushout square:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1.5,0) {$U(x)$};
\node (C) at (0,-1) {$c$};
\node (D) at (1.5,-1) {$c+_{a}U(x)$};
%\node (A') at (4,0) {$L(b)$};
%\node (B') at (6,0) {$x$};
%\node (C') at (4,-1) {$L(d)$};
%\node (D') at (6,-1) {$L(d)+_{L(b)}x$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i \circ h$} (B)
(A) edge node[left]{$k$} (C)
(B) edge node[right]{$$} (D)
(C) edge node[below]{$\psi$} (D);
%(A') edge node[above]{$I$} (B')
%(A') edge node[left]{$L(g)$} (C')
%(B') edge node[right]{$$} (D')
%(C') edge node[below]{$\theta$} (D');
\end{tikzpicture}
\]
exhibits $c+_a U(x)$ as a pushout of the span 
\[  c \xleftarrow{k} a \xrightarrow{i \circ h} U(x) 
\]
in $\A$. We may also take the first pushout square above and paste it with the original pushout square in $\A$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$b$};
\node (B) at (1.5,0) {$U(x)$};
\node (C) at (0,-1) {$d$};
\node (D) at (1.5,-1) {$d+_{b}U(x)$};
\node (E) at (-1.5,0) {$a$};
\node (F) at (-1.5,-1) {$c$};
\path[->,font=\scriptsize,>=angle 90]
(E) edge node [left] {$k$} (F)
(E) edge node [above] {$h$} (A)
(F) edge node [below] {$f$} (C)
(A) edge node[above]{$i$} (B)
(A) edge node[left]{$g$} (C)
(B) edge node[right]{$$} (D)
(C) edge node[below]{$\theta$} (D);
\end{tikzpicture}
\]
obtaining a pushout square 
which exhibits $d+_b U(x)$ as a pushout of the same span in $\A$.  The resulting natural isomorphism $\kappa$ between $c+_a U(x)$ and $d+_b U(x)$ then extends to a natural isomorphism between $z=\bot_c +_{\bot_a} x$ and $z'=\bot_d +_{\bot_b} x$ in $\X$:
\[
\begin{tikzpicture}[scale=1.5]
\node () at (5,0) {$\cong$};
\node (A) at (0,0) {$1$};

\node (A1) at (2,0.5) {$F(U(x))$};
\node (A2) at (4,0.5) {$F(0+_0 U(x))$};
\node (A3) at (6,0.5) {$F(c+_a U(x))$};

\node (B1) at (2,-0.5) {$F(U(x))$};
\node (B2) at (4,-0.5) {$F(0+_0 U(x))$};
\node (B3) at (6,-0.5) {$F(d+_b U(x))$};

\path[->,font=\scriptsize,>=angle 90]
(A) edge node[fill=white]{$x$} (A1)
(A) edge node[fill=white]{$x$} (B1)

(A1) edge node [above] {$F(\zeta)$} (A2)
(A2) edge node [above] {$F(!_c +_{!_a} 1)$} (A3)

(A1) edge node [fill=white] {$1$} (B1)
(A2) edge node [fill=white] {$1$} (B2)
(A3) edge node [fill=white] {$F(\kappa)$} (B3)

(B1) edge node [above] {$F(\zeta)$} (B2)
(B2) edge node [above] {$F(!_d +_{!_b} 1)$} (B3);

\end{tikzpicture}
\]
and consequently a natural isomorphism between our two objects of $(\A/U)_c$, showing that $U_1$ satisfies the Beck--Chevalley condition.
\end{proof}
